The problem and promise of scale in animal social multilayer networks

### 2020-02-10 paper/manuscript.Rmd
refs blah refs

Sociality exists across both temporal and spatial scales (Whitehead 2008). For
example, grooming requires close spatial proximity between conspecifics and
occurs over brief time periods [e.g. @Carter_2015], whereas social
association represents shared space use by members of the same social group
[Figure \@ref(fig:spacetime), @Franks_2010]. Further, for social
interactions or associations to occur individuals must share space, and thus
have overlapping home ranges. In the context of the social environment, home
range overlap is a relatively coarse scale of sociality across which individuals
share space over coarser spatial and temporal scales [e.g. @Piza_Roca_2018]. 
Although the social environment clearly scales spatially and temporally,
it remains unclear whether coarser scales of sociality, including social
association and home range overlap, vary predictably with social interaction
 [@Castles_2014; @Farine2015a]. Thus, we define social scale as the type of
social relationship observed that can be explicitly defined and measured [@Farine_2015a].

@@

Unlike traditional monolayer social networks, multilayer networks explicitly
consider social systems across contexts, including scale-dependent contexts
[@Pilosof_2017]. Multilayer networks are made up of multiple layers, each
representing (i) different classes of individuals, e.g. male or female; (ii) types 
of behaviours, e.g. grooming, travelling, or foraging; and (iii) spatial areas,
e.g. local or regional, or temporal windows, e.g. daily or seasonal 
 [@Kivela_2014; @Porter_2018]. Multilayer networks are relatively novel to studies of animal
behaviour [Silk_2018; @Finn_2019], although they have been used to
describe multidimensional human social systems, complex transportation networks,
and neural networks [@Silk_2018]. The role of scale within multilayer
social networks remains unexplored but provides a unique opportunity to develop
novel understanding of variation in the temporal or spatial scales associated
with social systems.

Our study is motivated by behavioural, landscape, and spatial ecology to build a
conceptual and analytical framework for scale-dependent multilayer networks and
we apply this framework to a case study of caribou (*Rangifer tarandus*) in
Newfoundland, Canada [@Peignier_2019]. We use this case study to address two
principal questions associated with scale-dependent multilayer networks. First,
to what degree does the social scale of relationships, spatial scale of
landscapes, and temporal scale of analysis influence our ability to interpret
complex social systems? Second, can variation in the social, spatial, and temporal
improve predictions associated with biological differences in seasonal resource
availability and social association? We conclude with perspectives for further
development and application of scale in multilayer networks with respect to 
phenotypes and individual fitness [@Webber_2018], movement ecology and collective
movement [@Jolles_2019], and habitat selection and space use <!-- TODO: (REF)-->.

@@

## Caribou socioecology
Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on social (visual and auditory perception), spatial (habitat
type), and temporal (seasonal). Caribou are gregarious ungulates with
fission-fusion dynamics [@Lesmerises_2018] that display temporal
[@Peignier_2019] and spatial (Webber et al. unpublished data) variation in
social networks. In winter, caribou dig holes in the snow, termed craters, to
access forage (Bergerud 1974)<!-- TODO: get Bergerud ref-->. As a result,
caribou tend to occupy and re-use craters once they are established and there is
considerably less access to forage than when the landscape is snow-free.
Importantly, craters exist on the landscape at multiple scales: the crater scale
(a single crater), the feeding area scale (multiple craters in close proximity),
and the winter range scale (all craters within an individual’s range)
[@Mayor_2009]. The distribution of forage for caribou is therefore seasonally
heterogeneous, with greater access in snow-free seasons compared to winter as
well as spatially heterogeneous in winter because the distribution of craters on
the landscape varies meaning that access to vegetation in winter is highly
variable for caribou. We therefore partitioned data into two discrete seasons:
summer (3 August - 20 September) and winter (1 January – 17 February) based on
caribou socioecology [@Peignier_2019].

## Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following [@Bjørneraas_2010]. 
We did not collar all female caribou in the herd; however, we assumed that our 
sample of collared animals was random.

>>>>

Sociality exists across both temporal and spatial scales (Whitehead 2008). For
example, grooming requires close spatial proximity between conspecifics and
occurs over brief time periods [e.g. @Carter_2015], whereas social
association represents shared space use by members of the same social group [Figure 1, @Franks_2009].
<!--Figure \@ref(fig:spacetime) TODO: something broken here?? --> 
Further, for social interactions or associations to occur individuals must share space, and thus
have overlapping home ranges. In the context of the social environment, home
range overlap is a relatively coarse scale of sociality across which individuals
share space over coarser spatial and temporal scales [e.g. @Piza_Roca_2018]. 
Although the social environment clearly scales spatially and temporally,
it remains unclear whether coarser scales of sociality, including social
association and home range overlap, vary predictably with social interaction
 [@Castles_2014; @Farine_2015a]. Thus, we define social scale as the type of
social relationship observed that can be explicitly defined and measured [@Farine_2015a]. 
<!-- Farine references comes out as D.R. Farine 2015 -->

@@

Unlike traditional monolayer social networks, multilayer networks explicitly
consider social systems across contexts, including scale-dependent contexts
[@Pilosof_2017]. Multilayer networks are made up of multiple layers, each
representing (i) different classes of individuals, e.g. male or female; (ii) types 
of behaviours, e.g. grooming, travelling, or foraging; and (iii) spatial areas,
e.g. local or regional, or temporal windows, e.g. daily or seasonal 
 [@Kivela_2014; @Porter_2018]. Multilayer networks are relatively novel to studies of animal
behaviour [@Silk_2018; @Finn_2019], although they have been used to
describe multidimensional human social systems, complex transportation networks,
and neural networks [@Silk_2018]. The role of scale within multilayer
social networks remains unexplored but provides a unique opportunity to develop
novel understanding of variation in the temporal or spatial scales associated
with social systems.

Our study is motivated by behavioural, landscape, and spatial ecology to build a
conceptual and analytical framework for scale-dependent multilayer networks and
we apply this framework to a case study of caribou (*Rangifer tarandus*) in
Newfoundland, Canada [@Peignier_2019]. We use this case study to address two
principal questions associated with scale-dependent multilayer networks. First,
to what degree does the social scale of relationships, spatial scale of
landscapes, and temporal scale of analysis influence our ability to interpret
complex social systems? Second, can variation in the social, spatial, and temporal
improve predictions associated with biological differences in seasonal resource
availability and social association? We conclude with perspectives for further
development and application of scale in multilayer networks with respect to 
phenotypes and individual fitness [@Webber_2018] <!-- TODO: ref should be 2018, but defaulting to 2017-->, 
movement ecology and collective
movement [@Jolles_2019], and habitat selection and space use. <!-- TODO: (REF)-->

@@

## Caribou socioecology
Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on social (visual and auditory perception), spatial (habitat
type), and temporal (seasonal). Caribou are gregarious ungulates with
fission-fusion dynamics [@Lesmerises_2018] that display temporal
[@Peignier_2019] and spatial (Webber et al. unpublished data) variation in
social networks. In winter, caribou dig holes in the snow, termed craters, to
access forage [@Bergerud_1974]. As a result, caribou tend to occupy and re-use 
craters once they are established and there is
considerably less access to forage than when the landscape is snow-free.
Importantly, craters exist on the landscape at multiple scales: the crater scale
(a single crater), the feeding area scale (multiple craters in close proximity),
and the winter range scale (all craters within an individual’s range)
[@Mayor_2009]. The distribution of forage for caribou is therefore seasonally
heterogeneous, with greater access in snow-free seasons compared to winter as
well as spatially heterogeneous in winter because the distribution of craters on
the landscape varies meaning that access to vegetation in winter is highly
variable for caribou. We therefore partitioned data into two discrete seasons:
summer (3 August - 20 September) and winter (1 January – 17 February) based on
caribou socioecology [@Peignier_2019].

## Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by [@Schaefer_2013]. Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following [@Bjørneraas_2010]. 
We did not collar all female caribou in the herd; however, we assumed that our 
sample of collared animals was random.
